An asymptotically normal out-of-sample test of equal predictive accuracy for nested models

### 2012-05-17 Paper.tex
Made several minor edits to the text of the main paper
- A few changes to formatting
- A few minor edits to the content

\documentclass[10pt,fleqn,draft]{article}

\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,graphicx,setspace,url,booktabs,tabularx,enumerate,slantsc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[nolists,nomarkers]{endfloat}
\usepackage[sort,round]{natbib}
\usepackage[top=1in,left=1.75in,right=1.75in,bottom=1.25in]{geometry}
\usepackage[small]{caption}
\newcolumntype{C}{>{\centering\arraybackslash}X}

\bibliographystyle{abbrvnat}
\newcommand\citepos[2][]{\citeauthor{#2}'s \citeyearpar[#1]{#2}}
\newcommand\poscw{\citeauthor{ClW:06}'s \citeyearpar{ClW:06,ClW:07}}
\newcommand\citen[1]{\citeauthor{#1}, \citeyear{#1}}
\frenchspacing
\input{tex/simulationdefinitions}
\input{tex/empiricsdefinitions}
\input{tex/empiricalresults}

@@

\newpage \noindent This paper proposes an out-of-sample (\oos) test
statistic that is asymptotically normal with mean zero even when the
models studied are nested, proves that common block bootstrap methods
consistently estimate the statistic's distribution, and shows how the
statistic can be used to study many models simultaneously while
maintaining strong control of the familywise error (\fwe),
i.e. ensuring that the probability that the hypothesis that the model
has equal predictive ability to a benchmark is incorrectly rejected
for any of the models is no higher than a prespecified level.  \oos\
tests are commonly used in International Macroeconomics,
Macroeconomics, and Finance \citep[see, for
example,][]{MeR:83,StW:03,GoW:08} and there is a substantial
literature developing the theoretical properties of these statistics,
beginning primarily with \citet{DiM:95} and
\citet{Wes:96}.\footnote{Other papers in this literature include
  \citet{WeM:98}, \citet{Mcc:98,Mcc:00},
  \citet{ClM:01,ClM:05-2,ClM:05,ClM:09b,ClM:11b,ClM:11},
  \citet{CCS:01}, \citet{CoS:02,CoS:04,CoS:07}, \citet{Whi:00},
  \citet{InK:04,InK:06}, \citet{Han:05}, \citet{Ros:05},
  \citet{ClW:06,ClW:07}, \citet{Ana:07}, \citet{GiR:09,GiR:10},
  \citet{HuW:10}, \citet{HLN:11}, \cite{InR:11}, \cite{Pin:11},
  \cite{RoS:11,RoS:11b}, and \citet{Cal:11}, among others.  For recent
  reviews of this literature and additional references, see
  \citet{McW:02}, \citet{CoS:06}, \citet{Wes:06}, \citet{ClM:11c},
  \citet{CoD:11}, and \citet{Gia:11}}

In a pair of papers, \citet{ClW:06,ClW:07} develop an \oos\ test of
the null hypothesis that a small benchmark model is correctly
specified.  Their test compares the forecasting performance of a pair
of nested models, and the null hypothesis is that the innovations in
the smaller model form a martingale difference sequence.  This test
procedure is popular, and one assumes that this is due in part to the
statistic's convenience.  The statistic is approximately normal after
adjusting for the estimation error of the larger model.  Normality
comes from a fixed-length rolling window, as in \citet{GiW:06}, and
the adjustment centers the statistic appropriately.  This statistic is
especially convenient because other \oos\ tests for similar hypotheses
\citep[among others]{Mcc:07,ClM:05,ClM:01,CoS:04,CoS:02,CCS:01} have a
nonstandard limit distribution and place restrictions on the models
under consideration, while other asymptotically normal statistics test
a different null hypothesis \citep{GiW:06} or place assumptions on the
models and \dgp\ that are often violated in empirical work
\citep{Wes:96,WeM:98,Mcc:00,DiM:95}.\footnote{\citet{DiM:95} assume
  that the models are not estimated. \citet{Wes:96,WeM:98,Mcc:00}
  implicitly assume that the models do not converge to the same limit,
  which rules out nesting.}

@@

Mixing window strategies is uncommon but needn't be. In most
applications, the null hypothesis imposes stability as well as equal
accuracy between the two models.  The benchmark model rarely allows
for breaks, parameter drift, or other forms of
instability,\footnote{Exceptions are \citepos{StW:07}
  \textsc{ima}(1,1) and \textsc{uc-sv} models of inflation.} but the
researcher is typically concerned about instability.  Indeed, concern
about instability is often given as a reason for doing an \oos\
analysis, especially with a short rolling window.\footnote{This
  motivation is discussed by \citet{StW:03}, \citet{PeT:05,PeT:07},
  \cite{GiW:06}, \citet{GoW:08}, \citet{ClM:09c}, and
  \cite{GiR:09,GiR:10}, among others.} A researcher could impose
stability on both models by using a recursive window or relax
stability for both by using a rolling window; either approach should
not affect the test's size, but may affect power.  But the researcher
could instead impose stability on the benchmark and relax it for the
alternative by using a recursive window for the benchmark and a
rolling window for the alternative model.  This approach could have a
power advantage and is similar in spirit to using a likelihood ratio
test instead of an \textsc{lm} or Wald test, which depend on just the
restricted or unrestricted model respectively.  Simulations presented
in Section~\ref{sec:2} show that mixing window strategies gives a
substantial power advantage when there are breaks.

This paper's statistic has a substantial advantage over existing \oos\
tests for nested models: the alternative can be essentially arbitrary,
as long as it obeys the necessary moment conditions.  In particular,
researchers can use model selection techniques like the \aic\ or \bic\
to determine the number of lags to include, the particular exogenous
variables to include, etc.  Other methods that test a similar
hypothesis are unable to handle these models \citep[except][which does
not allow the benchmark to be estimated]{ClW:06}; \citet{GiW:06} are
able to handle such models for both the alternative and the benchmark
but, as mentioned earlier, they test a different aspect of forecasting
performance.

@@

Since researchers often have a set of potential models and want to
know which of them significantly outperform the benchmark, procedures
that compare a single pair of models are of limited practical use.  As
\citet{Whi:00} demonstrates, looking at the naive $p$-values of the
individual tests is misleading, but researchers can get a valid test
by using the bootstrap to approximate the distribution of the largest
individual test statistic under the null (White calls this procedure
the ``bootstrap reality check'' or \brc).  This paper presents a test
for equal predictive ability of multiple models based on
\citepos{RoW:05} StepM, which uses a step-down procedure that
iteratively rejects models to achieve higher power than the \brc\ and
indicate which of the models improves on the benchmark (see
Theorem~\ref{res:2} for details).  Since this paper is concerned with
one-sided testing, I use the modified StepM procedure proposed by
\citet{Cal:11e}; the original StepM procedure unnecessarily sacrifices
power when testing one-sided hypotheses because, as \citet{Han:05}
points out for \citepos{Whi:00} \brc, alternative models that
substantially under-perform increase the test's critical values too
much.\footnote{Note that the modification suggested by \citet{Cal:11e}
  is different than \citepos{Han:05} and is explicitly designed for
  stepwise testing.}

This paper also presents a new result for the validity of block
bootstraps with \oos\ statistics, which is necessary to verify that
the StepM is valid.  Existing results on bootstrapping \oos\
statistics have some drawbacks.  \citet{Whi:00} and \citet{Han:05} use
the stationary bootstrap, but require the test sample to be much
smaller than the training sample, which is obviously inappropriate
here.  \citet{CoS:07} relax that requirement, but make the statistic
more complicated than necessary by adjusting the objective function of
the bootstrapped statistic to center it correctly.\footnote{Their
  recentering is required for consistency and does not give higher
  order accuracy.}  A parametric bootstrap, such as that used by
\cite{Mar:95}, \cite{Lut:99}, and \citet{ClM:11}, is an alternative
method, but requires the benchmark model to be correctly specified.
Although I focus on the null hypothesis that the benchmark is
correctly specified, this paper's block bootstrap methods remain valid
when the benchmark is misspecified and are relatively easy to
implement.  So this paper's result for bootstrapping \oos\ statistics
considerably improves on existing procedures.

@@

\section{Out-of-Sample Model Comparisons}\label{sec:1}
This section presents the new \oos\ statistic.  I'll present the
single-comparison statistic first and then the extension to multiple
comparisons.  Suppose for now that a researcher is interested in
predicting the target variable $y_{t+1}$ with a vector of regressors
$x_t$; further, let $v_t$ be another random process and suppose that
$(y_t, x_t, v_t)$ is stationary and weakly dependent
(Theorem~\ref{res:1} lists the detailed assumptions required).  In
addition, let $\beta_0 = (\E x_t x_t')^{-1} \E x_t y_{t+1}$ be the
pseudo-true coefficient of the regression of $y_{t+1}$ on $x_t$ and
define $\varepsilon_{t+1} = y_{t+1} - x_t'\beta_0$.  If the linear
model is correctly specified, so $\varepsilon_{t+1}$ is a martingale
difference sequence with respect to $\mathcal{F}_t = \sigma((x_t, v_t,
y_t), (x_{t-1}, v_{t-1}, y_{t-1}),\dots)$ then we can see immediately that
\begin{equation}
  \label{eq:4}
  \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \varepsilon_{t+1} (x_t'\beta_0
  - v_t) 
\end{equation}
with $R$ an arbitrary starting value\footnote{You will see momentarily
  why the summation begins at $R+1$ instead of $1$.} and $P = T - R$,
obeys an \mds\ \clt\ (assuming that its variance is uniformly
positive) and is asymptotically normal as $P \to \infty$.
Straightforward algebra \citep{ClW:07} shows that
\begin{equation}
  \label{eq:5}
  \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \varepsilon_{t+1} (x_t'\beta_0
  - v_t) = \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \Big[(y_{t+1} -
  x_t\beta_0)^2 + (y_{t+1} - v_t)^2 + (x_t'\beta_0 - v_t)^2 \Big].
\end{equation}

\citet{ClW:06,ClW:07} base their \oos\ statistic on the right side
of~\eqref{eq:5}, where $R$ is the size of the training sample and $P$
the size of the test sample used to evaluate the forecasting models.
\citet{ClW:06} use a second forecast $\hat{y}_{t+1}$ as $v_t$.  They
use a rolling window of length $R$ (i.e., $\hat{y}_{t+1}$ is a
function of $y_t, x_{t-1}, z_{t-1} \dots, y_{t-R+1}, x_{t-R}$ and
$z_{t-R}$ where $z_t$ is another weakly dependent random process),
which is kept finite as $T \to \infty$, so $\hat{y}_{t+1}$ inherits
the weak dependence properties of the variables used to estimate it.
Moreover, keeping $R$ finite ensures that the conditional variance
remains positive, so the \clt\ holds.  This method of ensuring
normality was introduced by \citet{GiW:06}.  The coefficients
$\beta_0$ are assumed to be zero so $\varepsilon_{t+1}$ is observed
directly (under the null hypothesis) and they propose using this
statistic to test the null that $\beta_0 = 0$.  In \citet{ClW:07},
$\beta_0$ is unknown but is estimated with a fixed-length rolling
window as well, so
\begin{equation*}
  \tilde{\beta}_t = \Big(\sum_{s=t-R+1}^t x_{t-1} x_{t-1}'\Big)^{-1}
  \sum_{s=t-R+1}^t x_{t-1} y_t
\end{equation*}
and $\hat{\varepsilon}_{t+1} = y_{t+1} - x_t'\tilde{\beta}_t$ replaces
$\varepsilon_{t+1}$ in the test statistic for $\beta_0 = 0$.
Unfortunately, $\hat{\varepsilon}_{t+1}$ is not an \mds, so the
statistic is no longer asymptotically mean-zero normal, even though
this approximation performs well in the simulations reported by
\citet{ClW:07}.

This paper proposes using the same basic \oos\ statistic, 
but using a recursive window to estimate $\beta_0$ and produce
$\hat{\varepsilon}_{t+1}$; i.e.
\begin{equation}
  \label{eq:8}
  \hat{\beta}_t = \Big(\sum_{s=2}^{t} x_{t-1} x_{t-1}'\Big)^{-1}
  \sum_{s=2}^t x_{t-1} y_t.
\end{equation}
\citepos{Wes:96} Theorem 4.1 implies that
\begin{equation*}
  \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \Big[(y_{t+1} -
  x_t\hat{\beta}_t)^2 + (y_{t+1} - v_t)^2 + (x_t'\hat{\beta}_t - v_t)^2 \Big]
\end{equation*}
is asymptotically normal for fairly arbitrary processes $v_t$, as long
as $v_t$ is weakly dependent and the \oos\ statistic has uniformly
positive variance.  Just as in \citet{ClW:06}, these conditions are
ensured if $v_t$ is another forecast of $y_{t+1}$ based on a fixed
length rolling window.  Theorem~\ref{res:1} presents the details of
this result.  The assumptions required are essentially the same in
existing papers (e.g. \citealp{Wes:96}; \citealp{WeM:98};
\citealp{Mcc:00}; \citealp{GiW:06}; and \citealp{ClW:06,ClW:07}).
Please see the original papers for a discussion of these assumptions.

\begin{thm}\label{res:1}
  Suppose that we have two models $\hat{y}_{0t}$ and $\hat{y}_{1t}$ to
  forecast the variable $y_t$, and have observations for
  $t=1,\dots,T+1$.  Assume the following hold:
  \begin{enumerate}
  \item \label{item:1} The benchmark forecast $\hat{y}_{0,t+1}$, is
    estimated using \ols\ with a recursive window: $\hat{y}_{0,t+1} =
    x_t'\hat{\beta}_t$ for some vector of predictors $x_t$ with
    $\hat{\beta}_t$ given by Equation~\eqref{eq:8}.  Also define
    $\beta_0 = \big(\sum_{t=2}^{T} \E x_{t-1} x_{t-1}'\big)^{-1}
    \sum_{t=2}^T \E x_{t-1} y_t$.
  \item \label{item:2} The alternative forecast, $\hat{y}_{1t}$, is
    estimated using a rolling window of fixed length $R$ (which is
    less than $T$), so $\hat{y}_{1,t+1} =
    \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})$ where $\psi$ is a known
    function and $z_t$ is a sequence of predictors that includes
    $x_t$.
  \item \label{item:3} The series $y_t$ and $z_t$ are \textsl{\ned} of size
    $-\frac12$ on a strong mixing series of size $-\frac{r}{r-2}$ for $r>2$ or a
    uniform mixing series of size $-\frac{r}{2r-2}$.  Also, $y_t$,
    $\hat{y}_{1t}$, and $x_t$ have uniformly bounded $2 r+\delta$
    moments for some positive $\delta$.
  \item \label{item:4} Define \[f_t(\beta) = (y_{t+1} - x_t'\beta)^2 -
    (y_{t+1} - \hat{y}_{1,t+1})^2 + (x_t'\beta - \hat{y}_{1,t+1})^2,\]
    $\bar{f}(\beta) = \tfrac1P \sum_{t=R+1}^T f_t(\beta)$, and $\bar f =
    \tfrac1P \sum_{t=R+1}^{T} f_t(\hat{\beta}_t)$, with $P = T - R$. The
    processes $f_t(\beta_0)$ and $x_t(y_{t+1} - x_t'\beta_0)$ are
    covariance stationary and $\bar f(\beta_0)$ has uniformly positive
    long run variance.
  \end{enumerate}
  Under the null hypothesis that $y_t - \hat{y}_{0t}(\beta_0)$ is a
  martingale difference sequence with respect to the filtration
  $\mathcal{F}_t = \sigma((y_t, z_{t}), (y_{t-1}, z_{t-1}),\dots)$,
  $\tfrac{\sqrt{P}}{\hat\sigma} \bar f \to^d N(0,1)$, where 
  \begin{align*}
    \hat{\sigma}^2 &= \hat{S}_{ff} + 2 \Pi (\hat{S}_{fg} + \hat{S}_{gg}), &
    \hat{S}_{ff} &= \tfrac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) - \bar
    f)^2, \\
    \hat{S}_{fg} &= \tfrac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) -
    \bar{f})(g_t(\hat{\beta}_t) - \bar g)', &
    \hat{S}_{gg} &= \tfrac1P \sum_{t=R+1}^T (g_t(\hat{\beta}_t) - \bar
    g)(g_t(\hat{\beta}_t) - \bar g)',
  \end{align*}
  $\Pi = 1 - \tfrac{R}{P} \log(1 + \tfrac{P}{R})$, $\bar{g} = \tfrac1T
  \sum_{t=R+1}^T g_t(\hat{\beta}_t)$, $\X' = [x_1,\dots,x_T]$, and
  \begin{equation*}
    g_t(\beta) = \Big\{\tfrac{2}{P}\sum_{t=R}^T x_t (2 x_t'\beta -
    y_{t+1} - \hat{y}_{1,t+1}) \Big\}' \big(\tfrac1T \X'\X \big)^{-1}
    x_t(y_{t+1} - x_t'\beta).
  \end{equation*}
\end{thm}

@@

\begin{rem}
  If one wants to test the less restrictive null hypothesis that
  $y_{t} - \hat{y}_{0t}$ is uncorrelated with $\hat{y}_{1t}$, one can
  replace $\hat{S}_{ff}$, $\hat{S}_{fg}$ and $\hat{S}_{gg}$ with their
  \hac\ counterparts.  Lemma~\ref{res:a5} presents a more general
  version of the theorem that can cover this case.
\end{rem}

@@

Note that the notation in this section is more general than that of
Section~\ref{sec:1}.  The assumptions required are generalizations of
the conditions of Theorems~\ref{res:1} and~\ref{res:2}.  See
\citet{Wes:96,Wes:06}, \citet{WeM:98}, and \citet{Mcc:00} for a
discussion of these conditions, as theirs are nearly identical.
Theorem~\ref{res:3} gives the result.

>>>>

\documentclass[12pt,fleqn]{article}

\usepackage{amsfonts}
\usepackage{amsmath,amsthm,amssymb,graphicx,setspace,url,booktabs,tabularx,enumerate,slantsc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[nolists,nomarkers]{endfloat}
\usepackage[sort,round,comma]{natbib}
\usepackage[top=1.25in,left=1.25in,right=1.25in,bottom=1.25in]{geometry}
\usepackage[small]{caption}
\newcolumntype{C}{>{\centering\arraybackslash}X}

\bibliographystyle{abbrvnat}
\newcommand\citepos[2][]{\citeauthor{#2}'s \citeyearpar[#1]{#2}}
\newcommand\poscw{\citeauthor{ClW:06}'s \citeyearpar{ClW:06,ClW:07}}
\newcommand\citen[1]{\citeauthor{#1}, \citeyear{#1}}
\frenchspacing\onehalfspacing
\input{tex/simulationdefinitions}
\input{tex/empiricsdefinitions}
\input{tex/empiricalresults}

@@

\newpage

\section{Introduction} This paper proposes an out-of-sample (\oos)
test statistic that is asymptotically normal with mean zero even when
the models studied are nested.  This paper also proves that common
block bootstrap methods consistently estimate the statistic's
distribution, and shows how the statistic can be used to study many
models simultaneously while maintaining strong control of the
familywise error (\fwe), i.e. ensuring that the probability that the
hypothesis that the model has equal predictive ability to a benchmark
is incorrectly rejected for any of the models is no higher than a
preset level.

\oos\ tests are commonly used in International Macroeconomics,
Macroeconomics, and Finance (see, for example, \citealt{MeR:83};
\citealt{StW:03}; and \citealt{GoW:08}) and there is a substantial
literature developing the theoretical properties of these statistics,
beginning primarily with \citet{DiM:95} and
\citet{Wes:96}.\footnote{Other papers in this literature include
  \citet{WeM:98}, \citet{Mcc:98,Mcc:00},
  \citet{ClM:01,ClM:05-2,ClM:05,ClM:11b,ClM:11,ClM:12},
  \citet{CCS:01}, \citet{CoS:02,CoS:04,CoS:07}, \citet{Whi:00},
  \citet{InK:04,InK:06}, \citet{Han:05}, \citet{Ros:05},
  \citet{ClW:06,ClW:07}, \citet{Ana:07}, \citet{GiR:09,GiR:10},
  \citet{HuW:10}, \citet{HLN:11}, \cite{InR:11}, \cite{Pin:11},
  \cite{RoS:11,RoS:11b}, and \citet{Cal:11}, among others.  For recent
  reviews of this literature and additional references, see
  \citet{McW:02}, \citet{CoS:06}, \citet{Wes:06}, \citet{ClM:11c},
  \citet{CoD:11}, and \citet{Gia:11}} In a pair of papers,
\citet{ClW:06,ClW:07} develop an \oos\ test of the null hypothesis
that a small benchmark model is correctly specified.  Their test
compares the forecasting performance of a pair of nested models, and
the null hypothesis is that the innovations in the smaller model form
a martingale difference sequence.  This test procedure is popular, and
one assumes that this is due in part to the statistic's convenience,
the statistic is approximately normal after adjusting for the
estimation error of the larger model.  Normality comes from a
fixed-length rolling window, as in \citet{GiW:06}, and the adjustment
centers the statistic appropriately.  This statistic is especially
convenient because other \oos\ tests for similar hypotheses
(\citealt{CCS:01}; \citealt{ClM:01,ClM:05}; \citealt{CoS:02,CoS:04};
and \citealt{Mcc:07}; among others) have a nonstandard limit
distribution and place restrictions on the models under consideration,
while other asymptotically normal statistics test a different null
hypothesis \citep{GiW:06} or place assumptions on the models and \dgp\
that are often violated in empirical work (\citealt{DiM:95};
\citealt{Wes:96}; \citealt{WeM:98};
\citealt{Mcc:00}).\footnote{\citet{DiM:95} assume that the models are
  not estimated. \citet{Wes:96}, \citet{WeM:98}, and \citet{Mcc:00}
  implicitly assume that the models do not converge to the same limit,
  which rules out nesting.}

@@

Mixing window strategies is uncommon but needn't be. In most
applications, the null hypothesis imposes stability as well as equal
accuracy between the two models.  The benchmark model rarely allows
for breaks, parameter drift, or other forms of
instability,\footnote{Exceptions are \citepos{StW:07}
  \textsc{ima}(1,1) and \textsc{uc-sv} models of inflation.} but the
researcher is typically concerned about instability.  Indeed, concern
about instability is often given as a reason for doing an \oos\
analysis, especially with a short rolling window.\footnote{This
  motivation is discussed by \citet{StW:03}, \citet{PeT:05,PeT:07},
  \cite{GiW:06}, \citet{GoW:08}, \citet{ClM:09c}, and
  \cite{GiR:09,GiR:10}, among others.} A researcher could impose
stability on both models by using a recursive window or relax
stability for both by using a rolling window; either approach should
not affect the test's size, but may affect power.  But the researcher
could instead impose stability on the benchmark and relax it for the
alternative by using a recursive window for the benchmark and a
rolling window for the alternative model.  This approach could have a
power advantage and is similar in spirit to using a Likelihood Ratio
Test instead of an \textsc{lm} or Wald test, which depend on just the
restricted or unrestricted model respectively.  Simulations presented
in Section~\ref{sec:2} show that this approach gives a substantial
power advantage when there are breaks.

This paper's statistic has a substantial advantage over existing \oos\
tests for nested models: the alternative can be essentially arbitrary
as long as high level moment conditions hold.  In particular,
researchers can use model selection techniques like the \aic\ or \bic\
to determine the number of lags to include, the particular exogenous
variables to include, etc.  Other methods that test a similar
hypothesis are unable to handle these models \citep[except][which does
not allow the benchmark to be estimated]{ClW:06}; \citet{GiW:06} are
able to handle such models for both the alternative and the benchmark
but, as mentioned earlier, they test a different aspect of forecasting
performance.

@@

Since researchers often have a set of potential models and want to
know which of them significantly outperform the benchmark, procedures
that compare a single pair of models are of limited practical value.
As \citet{Whi:00} demonstrates, looking at the naive $p$-values of
individual tests is misleading, but researchers can get a valid test
by using the bootstrap to approximate the distribution of the largest
individual test statistic under the null (White calls this procedure
the \textit{Bootstrap Reality Check} or \brc).  This paper presents a test
for equal predictive ability of multiple models based on
\citepos{RoW:05} StepM, which uses a step-down procedure that
iteratively rejects models to achieve higher power than the \brc\ and
indicate which of the models improves on the benchmark (see
Theorem~\ref{res:2} for details).

This paper also presents a new result for the validity of block
bootstraps with \oos\ statistics, which is necessary to verify that
the StepM is valid.  Existing results on bootstrapping \oos\
statistics have some drawbacks.  \citet{Whi:00} and \citet{Han:05} use
the stationary bootstrap, but require the test sample to be much
smaller than the training sample, which obviously does not hold here.
\citet{CoS:07} relax that requirement, but make the statistic more
complicated than necessary by adjusting the objective function of the
bootstrapped statistic to center it correctly.\footnote{Their
  recentering is required for consistency and does not imply higher
  order accuracy.}  A parametric bootstrap, such as that used by
\cite{Mar:95}, \cite{Lut:99}, and \citet{ClM:11}, is an alternative
method, but requires the benchmark model to be correctly specified.
Although I focus on the null hypothesis that the benchmark is
correctly specified, this paper's block bootstrap methods remain valid
when the benchmark is misspecified and are relatively easy to
implement.  So this paper's result for bootstrapping \oos\ statistics
considerably improves on existing procedures.

@@

\section{Out-of-Sample Model Comparisons}\label{sec:1}
This section presents the new \oos\ statistic.  I'll present the
single-comparison statistic first and then extend it to multiple
comparisons.  Suppose for now that a researcher is interested in
predicting the target variable $y_{t+1}$ with a vector of regressors
$x_t$; also let $v_t$ be another random process and suppose that
$(y_t, x_t, v_t)$ is stationary and weakly dependent
(Theorem~\ref{res:1} lists the assumptions formally).  In addition,
let $\beta_0 = (\E x_t x_t')^{-1} \E x_t y_{t+1}$ be the pseudo-true
coefficient of the regression of $y_{t+1}$ on $x_t$ and define
$\varepsilon_{t+1} = y_{t+1} - x_t'\beta_0$.  If the linear model is
correctly specified, so $\varepsilon_{t+1}$ is a martingale difference
sequence with respect to $\mathcal{F}_t \equiv \sigma((x_t, v_t, y_t),
(x_{t-1}, v_{t-1}, y_{t-1}),\dots)$, then we can see immediately that
\begin{equation}
  \label{eq:4}
  \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \varepsilon_{t+1} (v_t - x_t'\beta_0)
\end{equation}
obeys an \mds\ \clt\ (assuming that its variance is uniformly
positive) and is asymptotically normal as $P \to \infty$, with $R$ an
arbitrary starting value\footnote{It will be clear momentarily why the
  summation begins at $R+1$ instead of $1$.} and $P = T - R$.
Straightforward algebra \citep{ClW:07} shows that
\begin{equation}
  \label{eq:5}
  \tfrac{2}{\sqrt{P}} \sum_{t=R+1}^T \varepsilon_{t+1} (v_t -
  x_t'\beta_0) = \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \Big[(y_{t+1} - 
  x_t\beta_0)^2 - (y_{t+1} - v_t)^2 + (x_t'\beta_0 - v_t)^2 \Big].
\end{equation}

\citet{ClW:06,ClW:07} base their \oos\ statistic on the right side
of~\eqref{eq:5}, where $R$ is the size of the training sample and $P$
the size of the test sample used to evaluate the forecasting models.
\citet{ClW:06} use a second forecast $\hat{y}_{t+1}$ as $v_t$.  They
use a rolling window of length $R$ (i.e., $\hat{y}_{t+1}$ is a
function of $y_t, x_{t-1}, z_{t-1} \dots, y_{t-R+1}, x_{t-R}$ and
$z_{t-R}$ where $z_t$ is another weakly dependent random process),
which is kept finite as $T \to \infty$, so $\hat{y}_{t+1}$ inherits
the weak dependence properties of the variables used to estimate it.
Keeping $R$ finite ensures that the conditional variance remains
positive, so the sum obeys a \clt.  This method of ensuring normality
was introduced by \citet{GiW:06}.  The coefficients $\beta_0$ are
assumed to be zero so $\varepsilon_{t+1}$ is observed directly (under
the null hypothesis) and Clark and West propose using this statistic
to test the null that $\varepsilon_{t+1}$ is an \mds\ with respect to
$\mathcal{F}_t$.  In \citet{ClW:07}, $\beta_0$ is unknown but is
estimated with a fixed-length rolling window as well, so
\begin{equation*}
  \tilde{\beta}_t = \Big(\sum_{s=t-R+1}^t x_{t-1} x_{t-1}'\Big)^{-1}
  \sum_{s=t-R+1}^t x_{t-1} y_t
\end{equation*}
and $\hat{\varepsilon}_{t+1} = y_{t+1} - x_t'\tilde{\beta}_t$ replaces
$\varepsilon_{t+1}$ in the test statistic for \mds.  Unfortunately,
$\hat{\varepsilon}_{t+1}$ is not an \mds\ even when
$\varepsilon_{t+1}$ is, so the statistic is no longer asymptotically
mean-zero normal, even though this approximation performs well in the
simulations reported by \citet{ClW:07}.

This paper proposes using the same basic \oos\ statistic, 
but using a recursive window to estimate $\beta_0$ and produce
$\hat{\varepsilon}_{t+1}$; i.e.
\begin{equation}
  \label{eq:8}
  \hat{\beta}_t = \Big(\sum_{s=2}^{t} x_{t-1} x_{t-1}'\Big)^{-1}
  \sum_{s=2}^t x_{t-1} y_t.
\end{equation}
\citepos{Wes:96} Theorem 4.1 implies that
\begin{equation*}
  \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \Big[(y_{t+1} -
  x_t\hat{\beta}_t)^2 - (y_{t+1} - v_t)^2 + (x_t'\hat{\beta}_t - v_t)^2 \Big]
\end{equation*}
is asymptotically normal with mean zero under the null for fairly
arbitrary processes $v_t$, as long as $v_t$ is weakly dependent and
the \oos\ statistic has uniformly positive variance.  Just as in
\citet{ClW:06}, these conditions are ensured if $v_t$ is another
forecast of $y_{t+1}$ based on a fixed length rolling window.
Theorem~\ref{res:1} presents the details of this result.  The
assumptions required are essentially the same in existing papers
(e.g. \citealp{Wes:96}; \citealp{WeM:98}; \citealp{Mcc:00};
\citealp{GiW:06}; and \citealp{ClW:06,ClW:07}).  Please see the
original papers for a discussion of these assumptions.

\begin{thm}\label{res:1}
  Suppose that we have two models $\hat{y}_{0t}$ and $\hat{y}_{1t}$ to
  forecast the variable $y_t$, and have observations for
  $t=1,\dots,T+1$.  Assume the following hold:
  \begin{enumerate}
  \item \label{item:1} The benchmark forecast $\hat{y}_{0,t+1}$, is
    estimated using \textsl{\ols} with a recursive window:
    $\hat{y}_{0,t+1} = x_t'\hat{\beta}_t$ for some vector of
    predictors $x_t$ with $\hat{\beta}_t$ given by
    Equation~\eqref{eq:8}.  Also define $\beta_0 = (\E x_{t-1}
    x_{t-1}')^{-1} \E x_{t-1} y_t$ and assume that it does not depend
    on~$t$.
  \item \label{item:2} The alternative forecast, $\hat{y}_{1t}$, is
    estimated using a rolling window of fixed length $R$ (which is
    less than $T$), so $\hat{y}_{1,t+1} =
    \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})$ where $\psi$ is a known
    function and $z_t$ is a sequence of predictors that may include
    $x_t$.
  \item \label{item:3} The series $y_t$ and $z_t$ are \textsl{\ned} of size
    $-\frac12$ on a strong mixing series of size $-\frac{r}{r-2}$ for $r>2$ or a
    uniform mixing series of size $-\frac{r}{2r-2}$.  Also, $y_t$,
    $\hat{y}_{1t}$, and $x_t$ have uniformly bounded $2 r+\delta$
    moments for some positive $\delta$.
  \item \label{item:4} Define \[f_t(\beta) = (y_{t+1} - x_t'\beta)^2 -
    (y_{t+1} - \hat{y}_{1,t+1})^2 + (x_t'\beta - \hat{y}_{1,t+1})^2,\]
    $\bar{f}(\beta) = \tfrac1P \sum_{t=R+1}^T f_t(\beta)$, and $\bar f
    = \tfrac1P \sum_{t=R+1}^{T} f_t(\hat{\beta}_t)$, with $P = T -
    R$. The processes $f_t(\beta_0)$ and $x_t(y_{t+1} - x_t'\beta_0)$
    are covariance stationary and $\bar f(\beta_0)$ has uniformly
    positive and finite long run variance.
  \end{enumerate}
  Under the null hypothesis that $y_t - \hat{y}_{0t}(\beta_0)$ is a
  martingale difference sequence with respect to the filtration
  $\mathcal{F}_t = \sigma((y_t, z_{t}), (y_{t-1}, z_{t-1}),\dots)$,
  $\tfrac{\sqrt{P}}{\hat\sigma} \bar f \to^d N(0,1)$, where 
  \begin{align*}
    \hat{\sigma}^2 &= \hat{S}_{ff} + 2 \Pi (\hat{S}_{fg} + \hat{S}_{gg}), &
    \hat{S}_{ff} &= \tfrac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) - \bar
    f)^2, \\
    \hat{S}_{fg} &= \tfrac1P \sum_{t=R+1}^T (f_t(\hat{\beta}_t) -
    \bar{f})(g_t(\hat{\beta}_t) - \bar g)', &
    \hat{S}_{gg} &= \tfrac1P \sum_{t=R+1}^T (g_t(\hat{\beta}_t) - \bar
    g)(g_t(\hat{\beta}_t) - \bar g)',
  \end{align*}
  $\Pi = 1 - \tfrac{R}{P} \log(1 + \tfrac{P}{R})$, $\bar{g} = \tfrac1T
  \sum_{t=R+1}^T g_t(\hat{\beta}_t)$, $\X' = [x_1,\dots,x_T]$, and
  \begin{equation*}
    g_t(\beta) = \Big\{\tfrac{2}{P}\sum_{t=R}^T x_t (2 x_t'\beta -
    y_{t+1} - \hat{y}_{1,t+1}) \Big\}' \big(\tfrac1T \X'\X \big)^{-1}
    x_t(y_{t+1} - x_t'\beta).
  \end{equation*}
\end{thm}

@@

\begin{rem}
  If one wants to test the less restrictive null hypothesis that
  $y_{t} - \hat{y}_{0t}$ is uncorrelated with $\hat{y}_{1t}$ but not
  necessarily an \mds, one can replace $\hat{S}_{ff}$, $\hat{S}_{fg}$
  and $\hat{S}_{gg}$ with their \hac\ counterparts.
  Lemma~\ref{res:a5} presents a more general version of the theorem
  that can cover this case.
\end{rem}

@@

The notation in this section is more general than that of
Section~\ref{sec:1}.  The assumptions required are generalizations of
the conditions of Theorems~\ref{res:1} and~\ref{res:2}.  See
\citet{Wes:96,Wes:06}, \citet{WeM:98}, and \citet{Mcc:00} for a
discussion of these conditions, as theirs are nearly identical.
Theorem~\ref{res:3} gives the result.
